A Plasma Torus Around a Young Low-Mass Star

### 2025-03-19 paper/ms.tex
update

Figure~\ref{fig:hirescuts} shows a few regions of interest in the HIRES
spectra, which cover 3650-7960\,\AA.
Higher order Balmer lines including H$\gamma$ and H$\delta$ (the
$n=5\rightarrow2$ and $n=6\rightarrow2$ transitions respectively) do
show some variability outside the line core.
However, this variability is not clearly periodic in the
same manner as the emission seen in H$\alpha$.
This could be because there are insufficient hydrogen ions in the
relevant excited states, or due to the lower precision of the spectra
in these bluer regions.
The Ca HK doublet for instance is also detectable in emission, however
the continuum near it is not.
Chromospheric emission lines in the magnesium triplet order are
also detectable, but sufficiently blended to be rendered unusable.

@@

% TEMPLATE: IRAS041
%
\begin{table}
    \centering
    \begin{tabular}{lcr}
    \hline 
    \hline
    Parameter & Host & Source \\
    \hline 
    \multicolumn{3}{c}{Identifiers} \\
    \hline
    TIC & 141146667 & TESS \\
    Gaia & 860453786736413568 & Gaia\ DR3 \\
    %2MASS & J04154278+2909597 & J04154269+2909558 & 2MASS \\
    %ALLWISE & J041542.77+290959.5 & ... & ALLWISE\\
    \hline
    \multicolumn{3}{c}{Astrometry \& Radial Velocity} \\ 
    \hline
    $\alpha_{2000}$ & 11:05:15.09   & Gaia\ DR3 \\
    $\delta_{2000}$ & +59 15 05.57  & Gaia\ DR3 \\
    $\mu_{\alpha}$ (mas yr$^{-1}$ ) & -73.933 $\pm$ 0.022 & Gaia\ DR3 \\
    $\mu_{\delta}$ (mas yr$^{-1}$ ) &  32.262 $\pm$ 0.024 & Gaia\ DR3 \\
    $\pi$ (mas)                     &  17.324 $\pm$ 0.025 & Gaia\ DR3 \\
    RUWE                            &  1.23               & Gaia\ DR3 \\
    RV (\kms)                       & 0.61 $\pm$ 1.47     & HIRES \\
    \hline
    \multicolumn{3}{c}{Photometry} \\
    \hline
    $TESS$ (mag)                    & 13.283 $\pm$ 0.010 & TIC8\     \\
    $G$ (mag)                       & 14.701 $\pm$ 0.002 & Gaia\ DR3 \\
    $G_{\rm BP}$ (mag)              & 16.664 $\pm$ 0.008 & Gaia\ DR3 \\
    $G_{\rm RP}$ (mag)              & 13.398 $\pm$ 0.006 & Gaia\ DR3 \\
    $G_{\rm BP}$-$G_{\rm RP}$ (mag) &  3.276 $\pm$ 0.010 & Gaia\ DR3 \\
    $J$ (mag)                       & 11.401 $\pm$ 0.022 & 2MASS     \\
    $H$ (mag)                       & 10.793 $\pm$ 0.021 & 2MASS     \\
    $K_s$ (mag)                     & 10.473 $\pm$ 0.016 & 2MASS     \\
    $W1$ (mag)                      & 10.276 $\pm$ 0.023 & ALLWISE   \\ % Cutri+2013 II/328/allwise
    $W2$ (mag)                      & 10.070 $\pm$ 0.020 & ALLWISE   \\
    $W3$ (mag)                      &  9.838 $\pm$ 0.045 & ALLWISE   \\
    %$W4$ (mag)                      &  8.392 $\pm$ NOUNC & ALLWISE \\
	  %
    %TODO: eROSITA?  ROSAT is an emitter.  GALEX detection.
    % Ofek+2011 1.4Ghz emitter.  Helfand+2015 ditto.  Bruzewski+2021.
    %\hline
    %\multicolumn{3}{c}{Kinematics and Position} \\
    %\hline
    %$RV$ (\kms) & 0.61 $\pm$ 1.47 & HIRES \\
    %$U$ (\kms)  & & \\
    %$V$ (\kms)  & & \\
    %$W$ (\kms)  & & \\
    %$X$ (pc)    & -28.4 & \\
    %$Y$ (pc)    &  19.8 & \\
    %$Z$ (pc)    &  67.0 & \\
    \hline
    \multicolumn{3}{c}{Physical Properties} \\
    \hline
    $T_{eff}$ (K) & 2972 $\pm$ 40 & \cite{Bouma2024} SED fit\\
    $R_\star$ ($R_{\odot}$) & 0.42 $\pm$ 0.02 & \cite{Bouma2024} SED fit \\
    $P_{\rm rot}$ (hours) & $3.930\pm 0.001$ & TESS \\ 
    $v_{\rm eq}$ (\kms)  &  130$\pm$4  & Derived \\
    $v_{\rm eq} \sin i_\star$(km s$^{-1}$) & $138 \pm 8$ & HIRES \\
    $i_\star$($^\circ$) & $>$63 & Derived \\
    $A_V$ (mag) & 0 & \cite{Green2019} \\
    %$F_{bol}$ (erg cm$^{-2}$ s$^{-1}$ ) & todo & This work\\
    %$L_\star$ ($L_{\odot}$)  & todo & This work\\
    $M_\star$ ($M_{\odot}$)  & 0.22 $\pm$ 0.02  & PARSEC \cite{Chen2014}\\
    ${\rm EW}_{\rm Li}$ (m\AA) & $<$107 & HIRES (2$\sigma$)\\
    $t_{\rm CAMD}$ (Myr) & 30-150 &  Gaia DR3 \\
    $t_{\rm Li,emp}$ (Myr) & $>$20 &  HIRES, \cite{Jeffries2023} \\
    $t_{\rm Li,th}$ (Myr) & $>$35 &  HIRES, \cite{Feiden2016} \\
    $t_{\rm adopted}$ (Myr) & 35-150 &  -- \\
    \hline
    \end{tabular}
		\caption{Properties of \starname.  References:
    Gaia DR3 \cite{GaiaDR3}, TESS \cite{Ricker2015},
    TIC8 \cite{Stassun2019}, 2MASS \cite{Skrutskie2006}, ALLWISE
    \cite{Cutri2014}.}
    \label{tab:stparams}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{lcr}
  %\setlength{\tabcolsep}{2pt}
  \hline 
  \hline 
  Time [JD$_{\rm UTC}$] & RV (\kms) & $\sigma_{\rm RV}$ (\kms) \\
  % from tables/tab_TIC141146667_rel_rv.tex
  \hline 
  60357.450329 & 2.73 & 5.86 \\
  60357.461255 & -4.40 & 2.37 \\
  60357.472181 & -0.19 & 2.64 \\
  60357.483109 & 3.84 & 2.87 \\
  60357.494030 & 7.53 & 7.53 \\
  60357.504949 & -1.98 & 1.44 \\
  60357.515873 & 1.02 & 1.21 \\
  60357.526794 & 0.64 & 7.03 \\
  60357.537717 & -2.91 & 2.71 \\
  60357.548639 & 8.93 & 6.75 \\
  60357.559566 & 5.95 & 8.84 \\
  60357.570487 & -2.25 & 3.06 \\
  60357.581408 & 1.84 & 1.34 \\
  60357.592330 & 2.41 & 8.24 \\
  60357.603251 & -7.04 & 3.94 \\
  60357.614172 & -2.24 & 3.07 \\
  60357.625095 & -2.83 & 7.55 \\
  60357.636019 & -0.59 & 2.26 \\
  60357.646940 & 1.84 & 2.91 \\
  60357.657861 & 4.54 & 3.95 \\
  60357.668781 & 6.21 & 12.14 \\
  \hline
  \end{tabular}
  \caption{TIC~141146667 radial velocities relative to the systemic velocity.}
  \label{tab:rv}
\end{table}

>>>>

Figure~\ref{fig:hirescuts} shows a few regions of interest in the
HIRES spectra, which cover 3650-7960\,\AA.  Higher order Balmer lines
including H$\gamma$ and H$\delta$ (the $n=5\rightarrow2$ and
$n=6\rightarrow2$ transitions respectively) do show some variability
outside the line core.  However, this variability is not clearly
periodic in the same manner as the emission seen in H$\alpha$.  This
could be because there are insufficient hydrogen ions recombining from
the relevant excited states, or due to the lower precision of the
spectra in these bluer regions.  The Ca HK doublet for instance is
also detectable in emission, however the continuum near it is not.
Chromospheric emission lines in the magnesium triplet order are also
detectable, but sufficiently blended to be rendered unusable.

@@

% TEMPLATE: IRAS041
%
\begin{table}
    \centering
    \begin{tabular}{lcr}
    \hline 
    \hline
    Parameter & Host & Source \\
    \hline 
    \multicolumn{3}{c}{Identifiers} \\
    \hline
    TIC & 141146667 & TESS \\
    Gaia & 860453786736413568 & Gaia\ DR3 \\
    %2MASS & J04154278+2909597 & J04154269+2909558 & 2MASS \\
    %ALLWISE & J041542.77+290959.5 & ... & ALLWISE\\
    \hline
    \multicolumn{3}{c}{Astrometry \& Radial Velocity} \\ 
    \hline
    $\alpha_{2000}$ & 11:05:15.09   & Gaia\ DR3 \\
    $\delta_{2000}$ & +59 15 05.57  & Gaia\ DR3 \\
    $\mu_{\alpha}$ (mas yr$^{-1}$ ) & -73.933 $\pm$ 0.022 & Gaia\ DR3 \\
    $\mu_{\delta}$ (mas yr$^{-1}$ ) &  32.262 $\pm$ 0.024 & Gaia\ DR3 \\
    $\pi$ (mas)                     &  17.324 $\pm$ 0.025 & Gaia\ DR3 \\
    RUWE                            &  1.23               & Gaia\ DR3 \\
    RV (\kms)                       & 0.61 $\pm$ 1.47     & HIRES \\
    \hline
    \multicolumn{3}{c}{Photometry} \\
    \hline
    $TESS$ (mag)                    & 13.283 $\pm$ 0.010 & TIC8\     \\
    $G$ (mag)                       & 14.701 $\pm$ 0.002 & Gaia\ DR3 \\
    $G_{\rm BP}$ (mag)              & 16.664 $\pm$ 0.008 & Gaia\ DR3 \\
    $G_{\rm RP}$ (mag)              & 13.398 $\pm$ 0.006 & Gaia\ DR3 \\
    $G_{\rm BP}$-$G_{\rm RP}$ (mag) &  3.276 $\pm$ 0.010 & Gaia\ DR3 \\
    $J$ (mag)                       & 11.401 $\pm$ 0.022 & 2MASS     \\
    $H$ (mag)                       & 10.793 $\pm$ 0.021 & 2MASS     \\
    $K_s$ (mag)                     & 10.473 $\pm$ 0.016 & 2MASS     \\
    $W1$ (mag)                      & 10.276 $\pm$ 0.023 & ALLWISE   \\ % Cutri+2013 II/328/allwise
    $W2$ (mag)                      & 10.070 $\pm$ 0.020 & ALLWISE   \\
    $W3$ (mag)                      &  9.838 $\pm$ 0.045 & ALLWISE   \\
    %$W4$ (mag)                      &  8.392 $\pm$ NOUNC & ALLWISE \\
	  %
    % NOTE: ROSAT is an emitter.  GALEX detection.
    % Ofek+2011 1.4Ghz emitter.  Helfand+2015 ditto.  Bruzewski+2021.
    \hline
    \multicolumn{3}{c}{Physical Properties} \\
    \hline
    $T_{eff}$ (K) & 2972 $\pm$ 40 & \cite{Bouma2024} SED fit\\
    $R_\star$ ($R_{\odot}$) & 0.42 $\pm$ 0.02 & \cite{Bouma2024} SED fit \\
    $L_\star$ ($L_\odot$) & 0.0126 $\pm$ 0.0012 & \cite{Bouma2024} SED fit \\
    $P_{\rm rot}$ (hours) & $3.930\pm 0.001$ & TESS \\ 
    $v_{\rm eq}$ (\kms)  &  130$\pm$4  & Derived \\
    $v_{\rm eq} \sin i_\star$(km s$^{-1}$) & $138 \pm 8$ & HIRES \\
    $i_\star$($^\circ$) & $>$63 & Derived \\
    $A_V$ (mag) & 0 & \cite{Green2019} \\
    $M_\star$ ($M_{\odot}$)  & 0.22 $\pm$ 0.02  & PARSEC \cite{Chen2014}\\
    ${\rm EW}_{\rm Li}$ (m\AA) & $<$107 & HIRES (2$\sigma$)\\
    $t_{\rm CAMD}$ (Myr) & 30-150 &  Gaia DR3 \\
    $t_{\rm Li,emp}$ (Myr) & $>$20 &  HIRES, \cite{Jeffries2023} \\
    $t_{\rm Li,th}$ (Myr) & $>$35 &  HIRES, \cite{Feiden2016} \\
    $t_{\rm adopted}$ (Myr) & 35-150 &  -- \\
    \hline
    \end{tabular}
		\caption{Properties of \starname.  References:
    Gaia DR3 \cite{GaiaDR3}, TESS \cite{Ricker2015},
    TIC8 \cite{Stassun2019}, 2MASS \cite{Skrutskie2006}, ALLWISE
    \cite{Cutri2014}.}
    \label{tab:stparams}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{lcr}
  \hline 
  \hline 
  Time [BJD$_{\rm TDB}$] & RV (\kms) & $\sigma_{\rm RV}$ (\kms) \\
  % from tables/tab_TIC141146667_rel_rv.tex and
  % tables/tab_TIC141146667_times_rounded.csv
  \hline 
  2460357.954919 & 2.73 & 5.86 \\
  2460357.965845 & -4.40 & 2.37 \\
  2460357.976770 & -0.19 & 2.64 \\
  2460357.987698 & 3.84 & 2.87 \\
  2460357.998619 & 7.53 & 7.53 \\
  2460358.009538 & -1.98 & 1.44 \\
  2460358.020462 & 1.02 & 1.21 \\
  2460358.031383 & 0.64 & 7.03 \\
  2460358.042306 & -2.91 & 2.71 \\
  2460358.053228 & 8.93 & 6.75 \\
  2460358.064154 & 5.95 & 8.84 \\
  2460358.075075 & -2.25 & 3.06 \\
  2460358.085996 & 1.84 & 1.34 \\
  2460358.096918 & 2.41 & 8.24 \\
  2460358.107839 & -7.04 & 3.94 \\
  2460358.118760 & -2.24 & 3.07 \\
  2460358.129683 & -2.83 & 7.55 \\
  2460358.140606 & -0.59 & 2.26 \\
  2460358.151527 & 1.84 & 2.91 \\
  2460358.162448 & 4.54 & 3.95 \\
  2460358.173368 & 6.21 & 12.14 \\
  \hline
  \end{tabular}
  \caption{TIC~141146667 radial velocities relative to the systemic velocity.}
  \label{tab:rv}
\end{table}
